An Event-based Architecture for Multi-population Optimization Algorithms

### 2020-01-13 draft.tex
references fixed
no url, and no  list of editors, and no duplicates

Population-based algorithms are intrinsically parallel, and can be implemented to run asynchronously. % They needn't be intrinsically asynchronous, I think. Most are implemented synchronouslu - JJ  
%OK - M
The fitness of each individual can be evaluated independently of 
others, and similarly, each population could evolve in isolation. Since 
earlier works, researchers have been proposing some form of parallelization
\cite{muhlenbein1988evolution} to increase the scalability of these algorithms.
The island model was one of the first techniques proposed for parallelization,
which lead to an increased performance \cite{gorges1990explicit,grosso1985computer}. 
The concept was to divide a large population into communicated subpopulations. 
Since then, other population-based algorithms have adopted the concept, 
and researchers have found additional benefits
besides the execution speed; these include avoiding a premature convergence and
maintaining the diversity of the global population. Researchers use the term
multi-population based methods when generally referring to techniques using
subpopulations as part of their optimization strategy. % Some comment about asynchrony would be in order - JJ
% Comments below - Mario
Moreover, parallel implementations can run in two ways: synchronous and
asynchronous. When operations run in parallel but maintain synchronism with
other operations,  all operations need to finish to continue to the next step.
For instance, in a  parent/worker model, the parent needs to wait for all
workers to finish their operations before moving to the next iteration.  In
contrast, in an asynchronous execution,  operations are not synchronized with
each other. In the previous, the parent operation can continue to the next
iteration, even if a worker has not finished its work.  Several works have
proposed asynchronous algorithms \cite{coleman89,baugh2003asynchronous}, reporting benefits in execution time, and
scalability. In other works, sub-populations
are processed independently and communicate asynchronously through a central
repository \cite{sofea:cec2012,agajaj} or message queues \cite{salza2019speed, guervos2018introducing} .

@@

The parallel execution of multi-population algorithms has been the focus of many
works,  going from earlier hardware-based implementations using transputers \cite{gorges1990explicit},
multi-threaded \cite{merelo2019scaling}, and multi-core systems \cite{Serrano2008,lai2019adaptive}. % many citations here - JJ
Recently the trend has been on exploiting a higher number of processing units 
by using GPUs \cite{tan2015survey,li2007efficient}, or distributed systems; techniques used to achieve this include 
web-based \cite{JSON}, map-reduce \cite{fazenda2012},  grid \cite{munawar2010design,Gonzalez09}, 
voluntary computing \cite{MilkyWay}, and cloud computing 
\cite{GValdez2015,salza2019speed,valenzuela2015implementing,Sherry2012,FlexGP}. % many references here too - JJ
Cloud computing has become the standard way of running enterprise applications, not
only because of the convenience of the pay-as-you-go model or the non-existent
cost of administration, but also because it offers a way of describing the
infrastructure as part of the code, so that it is much easier to reproduce
results and this has been a boon for scientific computing.  However,  cloud computing
has also been evolving, going from simply putting in the cloud old-style monolithic applications, that is,
applications built on a single stack of services that communicate by layers, to
microservices \cite{microservices} and serverless architectures \cite{varghese2018next, Varghese2018849} that favor the parallel and
asynchronous communication of heterogeneous resources; in the process, 
a new process for designing {\em cloud native applications} has been created, 
with brand new methodologies and techniques \cite{Baldini2016287}. In these architectures,
services or even functions are seen as single processing points,  departing
from the monolithic and even distributed paradigms, become a loosely coupled collection of 
{\em stateless functions} \cite{malawski2017serverless}, that react to events and only come into existence while they 
are doing some processing. % many references more here - JJ

>>>>

Population-based algorithms are intrinsically parallel, and can be implemented to run asynchronously. % They needn't be intrinsically asynchronous, I think. Most are implemented synchronouslu - JJ  
%OK - M
The fitness of each individual can be evaluated independently of 
others, and similarly, each population could evolve in isolation. Since 
earlier works, researchers have been proposing some form of parallelization
\cite{muhlenbein1988evolution} to increase the scalability of these algorithms.
The island model was one of the first techniques proposed for parallelization,
which lead to an increased performance \cite{gorges1990explicit,grosso1985computer}. 
The concept was to divide a large population into communicated subpopulations. 
Since then, other population-based algorithms have adopted the concept, 
and researchers have found additional benefits
besides the execution speed; these include avoiding a premature convergence and
maintaining the diversity of the global population. Researchers use the term
multi-population based methods when generally referring to techniques using
subpopulations as part of their optimization strategy. % Some comment about asynchrony would be in order - JJ
% Comments below - Mario
Moreover, parallel implementations can run in two ways: synchronous and
asynchronous. When operations run in parallel but maintain synchronism with
other operations,  all operations need to finish to continue to the next step.
For instance, in a  parent/worker model, the parent needs to wait for all
workers to finish their operations before moving to the next iteration.  In
contrast, in an asynchronous execution,  operations are not synchronized with
each other. In the previous, the parent operation can continue to the next
iteration, even if a worker has not finished its work.  Several works have
proposed asynchronous algorithms \cite{coleman89,baugh2003asynchronous}, reporting benefits in execution time, and
scalability. In other works, sub-populations
are processed independently and communicate asynchronously through a central
repository \cite{sofea:cec2012,JSON} or message queues \cite{salza2019speed, guervos2018introducing} .

@@

The parallel execution of multi-population algorithms has been the focus of many
works,  going from earlier hardware-based implementations using transputers \cite{gorges1990explicit},
multi-threaded \cite{merelo2019scaling}, and multi-core systems \cite{Serrano2008,lai2019adaptive}. % many citations here - JJ
Recently the trend has been on exploiting a higher number of processing units 
by using GPUs \cite{tan2015survey,li2007efficient}, or distributed systems; techniques used to achieve this include 
web-based \cite{JSON}, map-reduce \cite{fazenda2012},  grid \cite{munawar2010design,Gonzalez09}, 
voluntary computing \cite{MilkyWay}, and cloud computing 
\cite{GValdez2015,salza2019speed,valenzuela2015implementing,FlexGP}. % many references here too - JJ
Cloud computing has become the standard way of running enterprise applications, not
only because of the convenience of the pay-as-you-go model or the non-existent
cost of administration, but also because it offers a way of describing the
infrastructure as part of the code, so that it is much easier to reproduce
results and this has been a boon for scientific computing.  However,  cloud computing
has also been evolving, going from simply putting in the cloud old-style monolithic applications, that is,
applications built on a single stack of services that communicate by layers, to
microservices \cite{microservices} and serverless architectures \cite{varghese2018next, Varghese2018849} that favor the parallel and
asynchronous communication of heterogeneous resources; in the process, 
a new process for designing {\em cloud native applications} has been created, 
with brand new methodologies and techniques \cite{Baldini2016287}. In these architectures,
services or even functions are seen as single processing points,  departing
from the monolithic and even distributed paradigms, become a loosely coupled collection of 
{\em stateless functions} \cite{malawski2017serverless}, that react to events and only come into existence while they 
are doing some processing. % many references more here - JJ
